Global Namespaces Scale as Well as Decoupled Namespaces

### 2017-02-21 paper/paper.tex
paper: re-organize methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{graphicx}

@@

@@

\subsection{Fault Tolerance}
\begin{figure*}[tb] \centering
\includegraphics[width=1\textwidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the MDSs. Each MDS maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure*}

@@

From this experiment we make 2 conclusions:
\begin{itemize}
  \item the metadata server is not overloaded because it can handle both workloads
  \item the metadata server needs to get global consistency from only 1 client
\end{itemize}

@@

\section{Methodology}

\subsection{Decoupled Namespaces in CephFS}

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications that can tolerate weaker consistency can decouple the
namespace, write updates to a local file, and delay metadata updates. The merge
and replay steps are optional if there is no need for a global namespace.
}\label{fig:decouple}
\end{figure}

%- How it works (timestamps)
Recall that CephFS uses RADOS (1) as a metadata store for all information about
files including the hierarchical namespace and (2) as a staging area for the
journal of updates before they are applied to the metadata store.  The journal
tool is used for disaster recovery and lets administrators view and modify the
journal of metadata updates; it can read the journal, erase events from the
journal, and apply the updates in the journal to the metadata store.  To apply
journal updates to the metadata store, the journal tool reads the journal
segments from RADOS objects and applies the update to the metadata store (which
are also stored as RADOS objects).

@@

@@

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them. There are two
types of objects:

1. journal objects (e.g., 200.000*)

2. metadata store objects (e.g., 1.inode)

We need to update the journal objects with the new events! How do we write out
new objects to RADOS (e.g., 200.000*)? These should be the corect sizes (e.g.,
4980 bytes) -- these log events that we are adding are not making them into
the journal objects... do we have to explicitly read them and then re-write
them?

TODO: pull the object, read it, append to it, make sure size changes. It looks
like we can only read these objects and apply doesn't actually touch them... it
only touches the metadata store objects (e.g., 1.inode*).

% TODO: figure out how import saves the journal on the metadata server
% - they aren't -- everything is stored in RADOS
% - we never touch the metadata servers!!!
% TODO: figure out how to read the header object
% TODO: figure out the mapping objects to journal segments (not 1x1)

@@

\subsubsection{Leveraging the Journal Tool}

%- How we leverage it
We leverage the journal tool's ability to reconstruct the namespace for the
metadata blob. First we scan journal and store all events in a map. Next we
create a new event and copy the root inode directory lump from another journal
event into the event's metadata blob. Finally, we add a new directory lump for
our new file or directory.

If decoupled namespaces are turned on, the client writes metadata updates
locally and merges the updates with the journal tool. The client that decouples
the namespace operates without any consistency and any conflicts at the merge
are resolved in favor of this client. Updates by other clients ({\it i.e.}
metadata writes to the global namespace) are overwritten. Cudele adopts the
following process when an application decouples the namespace:

\begin{enumerate}
  \item metadata server exports a snapshot of the file system to a binary file
  \item client materializes the snapshot in memory
  \item client operates on the data structures in memory
  \item client exports the new snapshot to a binary file
  \item metadata server merges the snapshot into the global namespace
\end{enumerate}

Step 3 is the most complicated and requires understanding how the snapshot is
materialized in memory.

\subsubsection{Operating on Snapshots}

Our first implementation attempted to re-create journal events using the same
libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
to instantiate a Ceph inode and directory entry for the current file/dir and
its parent.  This is too hard because there are too many moving parts in the
metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
traverse up and down namespace, etc.). So when I tried add dentries and inodes
it was trying to traverse up/down and it would almost always segfault when it
was looking for something. These metablobs are supposed to be self container --
the problem is I do not know what is supposed to go *inside* them.

Our second idea was to copy the metadata blog and change just what we needed.
For example, we would save a binary dump of a generic \texttt{mkdir} event on
disk. When the application makes a directory, this dump would be loaded and the
fields would be changed before being written back to disk. Rather than
traversing up and down a namespace in memory of a metadata server, we should
traverse up and down the namespace *inside* the metadata blob. This
implementation requires disk IO and editing the log event is non-trivial for
two reasons:

\begin{itemize}

  \item methods do not edit events; they just write them

  \item the metadata that the event touches (e.g., the metablob) is unorganized
  on disk for performance -- it is trade-off for writing data faster serially and
  reconstructing information slowly since failure is not considered the norm

\end{itemize}

Faced with these challenges we landed on our final implementation: load the
snapshot into the the data structures used to examine and replay journals, edit
those data structures, and write them out to disk as binary.

\section{Results}

In this section, we quantify the costs of consistency and fault tolerance in
CephFS. If the components that ensure these semantics (i.e. capabilities and
journals, respectively) can be mitigated or delayed, then global namespaces
perform as well as decoupled namespaces.

\textbf{Experimental Setup}: we run our experiments on a 9 OSD, 3 MDS, 1 MON
Ceph cluster. The clients use the Ceph kernel client, which has been in the
mainline Linux kernel since TODO. We use the kernel client so that we can find
the true create speed of the server; our experiments show a low CPU utilziation
for the clients which indicates that we are stressing the servers more. We also
turn caching off becuase, as shown in figureX there is little difference, in
terms of performance between caching and not caching when using the kernel
client.

>>>>

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{graphicx}
\usepackage{arydshln}

@@

\begin{tabular}{ r | l | l | l }
              & Decoupled & Global    & Implied   \\
              & Namespace & Namespace & Namespace \\\hline
  Example     & BatchFS & CephFS & PLFS \\
              & DeltaFS & IndexFS &     \\
  Consistency & eventual & strong & none? \\
  Fault Tol.  & node local & journal & journal \\
\end{tabular}

@@

In this section, we quantify the costs of consistency and fault tolerance in
CephFS. If the components that ensure these semantics (i.e. capabilities and
journals, respectively) can be mitigated or delayed, then global namespaces
perform as well as decoupled namespaces. We run our experiments on a 9 OSD, 3
MDS, 1 MON Ceph cluster. The clients use the Ceph kernel client, which has been
in the mainline Linux kernel since TODO. We use the kernel client so that we
can find the true create speed of the server; our experiments show a low CPU
utilziation for the clients which indicates that we are stressing the servers
more. We also turn caching off becuase, as shown in figureX there is little
difference, in terms of performance between caching and not caching when using
the kernel client.

\subsection{Fault Tolerance}
\label{sec:fault-tolerance}
\begin{figure*}[tb] \centering
\includegraphics[width=1\textwidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the MDSs. Each MDS maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure*}

@@

From this experiment we make 2 conclusions: (1) the metadata server is not
overloaded because it can handle both workloads and (2) the metadata server
needs to get global consistency from only 1 client

@@

\section{Methodology: Decoupled Namespaces}

\begin{figure}[tb]
\caption{Applications can decouple the namespace, write updates to a local
journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
these phases (represented by the arrows) can be combined to provide weaker
consistency or fault tolerance semantics.  }\label{fig:decouple}
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\end{figure}

In this section we describe Cudele, our prototype system that lets future
programmers compose mechanisms (Section~\S\ref{sec:cudeles-mechanisms}) to
provide the necessary guarantees
(Section~\S\ref{sec:setting-policies-with-cudele}) for their application.

\subsection{Cudele's Mechanisms}
\label{sec:cudeles-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace. The metadata store represents
the namespace as a tree of directory fragments and is easier to read/traverse.
On the other hand, the journal represent the namespace as a list of events. It
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more data to
read, it is unorganized, and many of the updates may be redundant.

% describe the mechanisms
Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Save, and
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Save" takes the in-memory journal and writes it to the client's
disk. ``Persist" saves the journal as a an object in the object store from the
client.

Next, we discuss how these mechanisms can be composed to get different
consistency and fault tolerance semantics.

%\begin{tabular}{ r | l }
%  \(\Rightarrow\)   & Description \\\hline
%  create            & events appended to in-memory journal \\
%  v.apply           & journal volatily applied to in-memory metadata store \\
%  save              & journal saved to client's disk \\
%  persist           & journal saved in object store \\
%  replay            & metadata servers materialize namespace \\
%  RPCs              & round trip remote procedure calls \\
%  stream            & metadata server streams journal into RADOS \\
%\end{tabular}

\subsection{Setting Policies with Cudele}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\caption{Cudele lets future programmers explore the spectrum of consistency and
fault tolerance semantics shown below.\label{table:spectrum}}
\begin{tabular}{ r | l | l | l }
  consistency     \(\rightarrow\) &&& \\  
  fault tolerance \(\downarrow\)  & none & eventual & global \\\hline
  none   & create   & create   & RPCs    \\
         &          & +v.apply &         \\\hdashline
  local  & create   & create   & RPCs    \\
         & +save    & +save    & +save   \\
         &          & +v.apply &         \\\hdashline
  global & create   & create   & RPCs    \\
         & +persist & +persist & +stream \\
         &          & +v.apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and fault tolerance guarantees that adminstrators
can construct is shown in Table~\ref{table:spectrum}. The columns are the
different consistency semantics and the rows cover the spectrum of fault
tolerance guantees. For consistency: ``none" means the system does not handle
merging updates into a global namespace and it is assumed that middleware or
the application manages consistency lazily; ``eventual" merges updates either
when the system has time (e.g., by a background daemon) or when the client is
done writing; and updates in ``global" consistency are seen immediately by all
clients. For fault tolerance, ``none" means that updates are volatile and will
be lost on a crash of any component. Stronger guarantees are made with
``local", which means updates will be retained if the client node recovers, and
``global", where all updates are always recoverable.

% which system they represent and which are impossible
The cells in Table~\ref{table:spectrum} encompass many exisiting storage
systems. POSIX systems like CephFS and IndexFS have global consistency and
fault tolerance; DeltaFS has consistency and fault tolerance set to ``none";
and BatchFS uses ``eventual" consistency and ``local" fault tolerance. These
are just a few of the HPC examples.

% how its done
To compose the mechanisms administrators inject which steps (described in
Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use in parallel
using a domain specific language. For example, to get the semantics of BatchFS,
the administrator would inject the following pipeline:\\

\noindent \texttt{create+save+v.apply}\\

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make much sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since global fault tolerance is stronger and has
more overhead than local fault tolerance.  Formulaically, valid steps
include:\\

\noindent \texttt{<create|RPCs> [+v.apply][+save][+persist][+stream]}

\subsection{Cudele as a Programmable Storage System}

% Programmability
We use a programmable storage approach to expose internal subsystem
abstractions as ``interfaces" for applications to create higher level
functionality and services.

\subsubsection{No Changes: Turning Knobs}

RPCs is the default behavior of the storage system. Depending on the cache
state and configuration each operation incurrs at least RPC. For example, if a
client or metadata server is not caching the directory inode, all creates
within that directory will result in a lookup and a create request. If the
directory inode is cached then only the create needs to be sent.

Re-using the debugging and testing features in the storage system, we can turn
``stream" on and off. If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run. Performance numbers for enabling and disabling
the stream phase are presented back in Section~\S\ref{sec:fault-tolerance}.

\subsubsection{External Library: Reads/Writes Journal Events}

% How it works
For ``create", ``save", and ``persist" Cudele provides a library for clients to
link into.  Recall that CephFS uses the object store (1) as a metadata store
for all information about files including the hierarchical namespace and (2) as
a staging area for the journal of updates before they are applied to the
metadata store. As discussed previously, the metadata store is optimized for
reading while the journal is optimized for writing.  Cudele re-uses the journal
tool to read/write log events to memory and persistent storage.

% creates
The journal tool is used for disaster recovery and lets administrators view and
modify the journal of metadata updates; it can read the journal, erase events
from the journal, and apply the updates in the journal to the metadata store.
To apply journal updates to the metadata store, the journal tool reads the
journal segments from object store objects and applies the update to the
metadata store (which are also stored as object store objects).  For create,
clients write to an in-memory journal and updates are merged into the global
namespace by replaying them onto the metadata store in the object store.

% persist/save
For save, clients write serialized log events to a file on local disk and for
persist, clients store the file in the objects store. The overheads for save
and persist is the write bandwidth of the local disk and object store,
respectively.

% Level of intrusion
This required no changes to Ceph because the metadata server knows how to read
the events we were writing into the object store.  The client writes metadata
updates locally and merges the updates with the journal tool. The client that
decouples the namespace operates without any consistency and any conflicts at
the merge are resolved in favor of this client. Updates by other clients ({\it
i.e.} metadata writes to the global namespace) are overwritten.  We leverage
the journal tool's ability to reconstruct metadata events in memory. The client
library is shown in Figure~\ref{fig:decouple}.  Cudele adopts the following
process when an application decouples the namespace:

\begin{enumerate}
  \item ``\textbf{decoupled}": metadata server exports the journal events to a file
  \item ``\textbf{transfer}": the file is pulled by the client from the metadata server
  \item ``\textbf{snapshot}": client reads the file and materializes a snapshot of the
        namespace in memory
\end{enumerate}

% Why we re-use stuff
By using re-using the journal subsystem to implement the namespace decoupling,
Cudele leverages the write/read optimized data structures, the formats for
persisting events (similar to TableFS's SSTables), and the functions for
replaying events onto the internal namespace data structures.

%Step 3 is the most complicated and requires understanding how the snapshot is
%materialized in memory. 
%
%\subsubsection{Operating on Snapshots} 
%
%Our first implementation attempted to re-create journal events using the same
%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
%to instantiate a Ceph inode and directory entry for the current file/dir and
%its parent.  This is too hard because there are too many moving parts in the
%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
%traverse up and down namespace, etc.). So when I tried add dentries and inodes
%it was trying to traverse up/down and it would almost always segfault when it
%was looking for something. These metablobs are supposed to be self container --
%the problem is I do not know what is supposed to go *inside* them. 
%
%Our second idea was to copy the metadata blog and change just what we needed.
%For example, we would save a binary dump of a generic \texttt{mkdir} event on
%disk. When the application makes a directory, this dump would be loaded and the
%fields would be changed before being written back to disk. Rather than
%traversing up and down a namespace in memory of a metadata server, we should
%traverse up and down the namespace *inside* the metadata blob. This
%implementation requires disk IO and editing the log event is non-trivial for
%two reasons:
%
%\begin{itemize}
%
%  \item methods do not edit events; they just write them
%
%  \item the metadata that the event touches (e.g., the metablob) is unorganized
%  on disk for performance -- it is trade-off for writing data faster serially and
%  reconstructing information slowly since failure is not considered the norm
%
%\end{itemize}
%
%Faced with these challenges we landed on our final implementation: load the
%snapshot into the the data structures used to examine and replay journals, edit
%those data structures, and write them out to disk as binary.

\subsubsection{Storage System Changes: Volatile Merge}

%The metadata objects are located with naming schemes (200.000* for journal
%objects and 1.inode for metadata storage objects).

% How it works: socket for changing daemon's internal state (debugging, logging, behaviour)
% 1. API for putting state into the daemon dynamically
% 2. Documentation all the tunables
% 3. Hooks directly into daemon code so we can use any parsing functionality in there

\section{Implementation}
\label{sec:implementation}

% What we actually implemented
Of these 6 mechanisms only 4 had to implemented. ``RPCs" and ``Stream" can be
achieved with tunables in configuration files for the storage system (e.g., metadata cache size,
logging on/off, etc.). 
``Persist", ``Save", and ``Create" are implemented in
separate programs and libraries and do not require modifications to the storage
system. ``Volatile Apply" requires changes to the metadata server to inject
updates back into the global namespace.

% Outline of the section
In this section, we describe how we decoupled the consistency and fault
tolerance policies from their mechanisms to allow applications to choose the
semantics they need. Each section corresponds ot the labeled arrows in
Figure~\ref{fig:decouple}.

\subsection{RPCs (Existing CephFS Implementation)}

CephFS uses round-trip remote procedure calls (RPCs) for most of

\subsection{Stream (Existing Implementation)}

@@

\subsection{Create}

@@

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them.

@@

%- How we leverage it

\subsection{Volatile Apply (v.apply)}
\label{sec:vapply}
\subsection{Save}
\label{sec:save}
\subsection{Persist}
\label{sec:persist}

\section{Results}
\subsection{Microbenchmarks}

\subsubsection{Per phase latencies}
\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/global-v-decoupled.png}
\caption{Compared to the \texttt{create} phase, saving and persisting
updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
slower.  The disadvantage of decoupling the namespace is the merge phase where
updates are applied to the metadata store (\texttt{create+apply}, resulting in a
905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
\end{figure}
